---
title: "practical machines"
author: "Binayak"
date: "17/02/2021"
output:
  pdf_document: default
  html_document: default
---
Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement ??? a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset)

###Data
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

### Library Loading
```{r,echo= TRUE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(corrplot)
library(gbm)
```
### Data Loading
```{r,echo=TRUE}
train1<-read.csv("C:/Users/binayak mishra/Downloads/pml-training.csv",header = TRUE)
test1<-read.csv("C:/Users/binayak mishra/Downloads/pml-testing.csv",header = TRUE)
```
### Data Cleaning
```{r,echo=TRUE}
training<-train1[,colSums(is.na(train1))==0]
testing<-test1[,colSums(is.na(test1))==0]
training<-training[,-c(1:7)]
testing<-testing[,-c(1:7)]
```
### Dataset for prediction
```{r,echo=TRUE}
set.seed(12345)
intrain<-createDataPartition(training$classe,p =0.7,list = FALSE)
training<-training[intrain,]
testing<-training[-intrain,]
### Cleaning by removing nearly zero variance variable
Nz<-nearZeroVar(training)
training<-training[,-Nz]
testing<-testing[,-Nz]
```
### Plotting the correlation Matrix
```{r,echo=TRUE}
cormat<-cor(training[,-53])
corrplot(cormat,order = "FPC",method = "color",type = "upper",tl.cex = 0.8,tl.col = rgb(0,0,0))
highcorr<-findCorrelation(cormat,cutoff = 0.75)
names(training)[highcorr]
```
### MODEL BUILDING USING CLASSIFICATION TREE
```{r,echo= TRUE}
Mod1Tr<-rpart(classe ~ .,data = training, method = "class")
fancyRpartPlot(Mod1Tr)
predmod1Tr<-predict(Mod1Tr,testing,type = "class")
classes<- as.factor(testing$classe)
cmTr<-confusionMatrix(predmod1Tr,classes)
cmTr
```
### MODEL BUILDING USING RANDOM FOREST
```{r,echo=TRUE}
contRf<-trainControl(method ="cv",number=5)
contRf
MODRF11<- train(classe~., data = training,method="rf",trControl=contRf,verbose= FALSE)
predictF1<-predict(MODRF11,newdata = testing)
```
```{r,echo=TRUE}
classes<- as.factor(testing$classe)
cmrf1<-confusionMatrix(predictF1,classes)
cmrf1
### PLOTTING THE Random forest model
plot(MODRF11)
plot(cmrf1$table,col=cmrf1$byClass,main= paste("Random forest Confusion Matrix: Accuracy=",round(cmrf1$overall["Accuracy"],4)))
```
### MODEL BUILDING USING GBM
```{r,echo= TRUE}
set.seed(12345)
controlGBM<-trainControl(method = "repeatedcv",number = 5,repeats = 1)
modgbm<-train(classe ~.,data = training, method="gbm", trControl=controlGBM, verbose= FALSE)
modgbm$finalModel
print(modgbm)
predictGBM<-predict(modgbm, newdata = testing)
cmGBM<- confusionMatrix(predictGBM,classes)
cmGBM
``` 
### Result
We use random forest algorithm to predict the test set
```{r,echo= TRUE}
FinalResult<-predict(MODRF11,newdata = test1)
FinalResult
```
###  Conclusion
As can be seen that random forest gives us a prediction accuracy of 100% which is better than gbm(97.5% accuracy) and decision tree whose accuracy is (75%) 



